---
title: "CSIAPPS REST API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CSIAPPS REST API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(csiapps)
```

### Preamble

The following environment variables must be set to make requests to the `CSIAPPS` REST API:

-   `CSIAPPS_ACCESS_TOKEN`: The access token for authenticating API requests. This token should have the necessary permissions to access the desired endpoints.

### Overview

`csiapps` includes a generic `make_request()` function to make requests to the `CSIAPPS` REST API.
Please refer to the [CSIAPPS Swagger documentation](https://apps.csiontario.ca/api/swagger/)
for full details on available endpoints, parameters, and response formats. 

Below are some common use cases for interacting with the API endpoints using `csiapps`.

### Authorization Status

The most basic endpoint is the `/api/csiauth/me/` endpoint, which can be used to check the
status of the user making the request. This endpoint requires authentication, so if the request is successful, it indicates that your access token is valid and has the necessary permissions to access the REST API.

```{r}
result <- make_request(
  endpoint = "api/csiauth/me/"
)
```


### Data Warehouse Ingestion

To ingest raw data into the warehouse, we make a POST request to the `/api/warehouse/ingestion/primary/` endpoint.
Users supply the `source` (uuid) of the target table,
a list of `records` to be ingested, and (optionally) the `subject_field` by which the record can be uniquely identified.
The records must comply with the schema defined for the target table.

```{r}
result <- make_request(
  endpoint = "api/warehouse/ingestion/primary/",
  method = "POST",
  body = list(
    source = Sys.getenv("SOURCE_UUID"),
    records = list(
      list(id = "xxxx", ...),
      list(id = "yyyy", ...),
      ...
    ),
    subject_field = "id"
  )
)
```

### Data Record Retrieval 

To retrieve data records from the warehouse, we make an API request to the `/api/warehouse/data-records/` endpoint.
The only required parameter is the `source_uuid` of the requested table, but there are several optional parameters for additional filtering. Note that we can set the `paginate` argument to `TRUE` to retrieve all records matching the query parameters.

```{r list-records}
records <- make_request(
    endpoint = "api/warehouse/data-records",
    query = list(source_uuid = Sys.getenv("SOURCE_UUID")),
    paginate = T
    )
```

<!-- ### Data Record Flattening -->

<!-- After retrieving records, `flatten_record()` can be used to flatten nested records into a more tabular format for easier analysis. It retains the original `data` payload and also includes additional metadata fields such as the unique `id`, the timestamp of when the record was collected, and any other relevant information. -->

<!-- ```{r flatten-records} -->
<!-- flat_records <-  -->
<!--   records |>  -->
<!--   dplyr::pull(results) |>  -->
<!--   lapply(flatten_record) -->
<!-- ``` -->

