---
title: "Introduction to CSIAPPS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to CSIAPPS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```


## Developing Shiny Web Applications

### Example
Suppose that you have the following shiny web application:

```{r}
library(shiny)

df = faithful[, 2]

ui <- fluidPage(

    titlePanel("Old Faithful Geyser Data"),

    sidebarLayout(
        sidebarPanel(
            sliderInput("bins",
                        "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30)
        ),

        mainPanel(
           plotOutput("distPlot")
        )
    )
)

server <- function(input, output) {

    output$distPlot <- renderPlot({
        bins <- seq(min(df), max(df), length.out = input$bins + 1)

        hist(df, breaks = bins, col = 'darkgray', border = 'white',
             xlab = 'Waiting time to next eruption (in mins)',
             main = 'Histogram of waiting times')
    })
}

shinyApp(ui = ui, server = server)
```

### Process
To migrate this app within the `CSIAPPS` ecosystem, we can leverage several functions provided by `csiapps`.

1.  We specify which institute internal API calls should be made for (such as authentication redirects) by the application using `set_institute()`.

```{r set-institute}
# institute can be set to one of "csiontario" or "csipacific"
csiapps::set_institute("csiontario")
```

2.  We then run `check_secrets()` to ensure that all environment variables required have been set. The `verbose` argument, which is `FALSE` by default, can be set to `TRUE` to print out the values of the environment variables that are being checked. `check_secrets()` will throw an error if any of the required environment variables are not present, making it useful for debugging.

```{r check-secrets}
csiapps::check_secrets(verbose = FALSE)
```

3.  In the application, for code defined outside but used within the `server` function, we use
`global_wrapper()` to ensure that it is accessible by internal `csiapps` functions.

```{r global-wrapper}
csiapps::global_wrapper({
  df = faithful[, 2]
})
```

4.  For the `ui`, it can simply be wrapped within `ui_wrapper()`, which includes additional 
code to redirect the app to CSIAPPS for user authentication and aesthetic formatting.
The same principle applies to the `server` function, which can be wrapped within `server_wrapper()`.

```{r ui-wrapper}
# original code

ui <- ...

server <- ...

shinyApp(ui = ui, server = server)

# wrapped code

shinyApp(ui = csiapps::ui_wrapper(ui), server = csiapps::server_wrapper(server))
```

Thefore, the full code for the app, after migration, would look like this:

```{r}
library(shiny)
library(csiapps)

# CSIAPPS Setup
set_institute("csiontario")
check_secrets()

global_wrapper({
  df = faithful[, 2]
})

# Define UI for application that draws a histogram
ui <- fluidPage(

    titlePanel("Old Faithful Geyser Data"),

    sidebarLayout(
        sidebarPanel(
            sliderInput("bins",
                        "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30)
        ),

        mainPanel(
           plotOutput("distPlot")
        )
    )
)

# Define server logic required to draw a histogram
server <- function(input, output) {

    output$distPlot <- renderPlot({
        bins <- seq(min(df), max(df), length.out = input$bins + 1)

        # draw the histogram with the specified number of bins
        hist(df, breaks = bins, col = 'darkgray', border = 'white',
             xlab = 'Waiting time to next eruption (in mins)',
             main = 'Histogram of waiting times')
    })
}

# Run the application 
shinyApp(ui = ui_wrapper(ui), server = server_wrapper(server))
```

## Interfacing with the CSIAPPS REST API

`csiapps` includes several functions to make API calls to the `CSIAPPS` data warehouse
while abstracting away the underlying HTTP requests. 

We first generate a warehouse client object using `new_warehouse_client()`.

```{r warehouse-client}
wc <- csiapps::new_warehouse_client()
```

### `ingest_raw()`

To ingest raw data into the warehouse, `ingest_raw()`, which is a convenience wrapper for the `/api/warehouse/ingestion/primary/` endpoint, can be used. It requires the `SOURCE_UUID` of the target table,
a list of `records` to be ingested, the `subject_field` by which the record can be uniquely identified.
The records must comply with the schema defined for the target table.

```{r}
wc$ingest_raw(
  source_uuid = Sys.getenv("SOURCE_UUID"),
  records = list(
    list(id = "xxxx", ...),
    list(id = "yyyy", ...),
    ...
  ),
  subject_field = "id"
)
```

### `list_records()`

To retrieve data from the warehouse, `list_records()`,
a convenience wrapper for the `/api/warehouse/data-records/` endpoint, can be used.
It requires the `SOURCE_UUID` of the requested table and also includes
several optional parameters for filtering and pagination. 

```{r list-records}
records <- wc$list_records(
  source_uuid = "SOURCE_UUID", 
  collected_after = Sys.Date() - 30,
  collected_before = Sys.Date(),
  extra_params = list(role = "primary")
  )
```

After retrieving records, `flatten_record()` can be used to flatten nested records into a more tabular format for easier analysis. It retains the original `data` payload and also includes additional metadata fields such as the unique `id`, the timestamp of when the record was collected, and any other relevant information.

```{r flatten-records}
flat_records <- lapply(records, flatten_record)
```

To use native HTTP requests for more customized interactions with the API, please refer to the 
Swagger documentation, which can be found at https://apps.csiontario.ca/api/swagger/.
It includes details on the available endpoints, request parameters, and response formats.
