---
title: "Developing Shiny Web Applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Developing Shiny Web Applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Example
Suppose that you have the following shiny web application:

```{r}
library(shiny)

df = faithful[, 2]

ui <- fluidPage(

    titlePanel("Old Faithful Geyser Data"),

    sidebarLayout(
        sidebarPanel(
            sliderInput("bins",
                        "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30)
        ),

        mainPanel(
           plotOutput("distPlot")
        )
    )
)

server <- function(input, output) {

    output$distPlot <- renderPlot({
        bins <- seq(min(df), max(df), length.out = input$bins + 1)

        hist(df, breaks = bins, col = 'darkgray', border = 'white',
             xlab = 'Waiting time to next eruption (in mins)',
             main = 'Histogram of waiting times')
    })
}

shinyApp(ui = ui, server = server)
```


To migrate this app within the `CSIAPPS` ecosystem, we can leverage several functions provided by `csiapps`.

### 1. `set_institute()`

We specify which institute internal API calls should be made for (such as authentication redirects) by the application using `set_institute()`.

```{r set-institute}
# institute can be set to one of "csiontario" or "csipacific"
csiapps::set_institute("csiontario")
```

### 2. `check_secrets()`

We then run `check_secrets()` to ensure that all environment variables required have been set. The `verbose` argument, which is `FALSE` by default, can be set to `TRUE` to print out the values of the environment variables that are being checked. `check_secrets()` will throw an error if any of the required environment variables are not present, making it useful for debugging.

```{r check-secrets}
csiapps::check_secrets(verbose = FALSE)
```

### 3. `global_wrapper()`

For code defined outside but used within the `server` function, we use
`global_wrapper()` to ensure that it is accessible by internal helper functions.

```{r global-wrapper}
csiapps::global_wrapper({
  df = faithful[, 2]
})
```

### 4. `ui_wrapper()` and `server_wrapper()`

The `ui` page and `server` function can simply be wrapped by `ui_wrapper()` and `server_wrapper()`, respectively.
These convenience functions include additional 
code to redirect the application to CSIAPPS for user authentication and provide aesthetic formatting.

```{r ui-wrapper}
# original code

ui <- ...

server <- ...

shinyApp(ui = ui, server = server)

# wrapped code

shinyApp(ui = csiapps::ui_wrapper(ui), server = csiapps::server_wrapper(server))
```

### Summary

Therefore, the full code for the app, after migration, would look like this:

```{r}
library(shiny)
library(csiapps)

# CSIAPPS Setup
set_institute("csiontario")
check_secrets()

global_wrapper({
  df = faithful[, 2]
})

# Define UI for application that draws a histogram
ui <- fluidPage(

    titlePanel("Old Faithful Geyser Data"),

    sidebarLayout(
        sidebarPanel(
            sliderInput("bins",
                        "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30)
        ),

        mainPanel(
           plotOutput("distPlot")
        )
    )
)

# Define server logic required to draw a histogram
server <- function(input, output) {

    output$distPlot <- renderPlot({
        bins <- seq(min(df), max(df), length.out = input$bins + 1)

        # draw the histogram with the specified number of bins
        hist(df, breaks = bins, col = 'darkgray', border = 'white',
             xlab = 'Waiting time to next eruption (in mins)',
             main = 'Histogram of waiting times')
    })
}

# Run the application 
shinyApp(ui = ui_wrapper(ui), server = server_wrapper(server))
```
